First steps towards a formalization of Forcing

### 2018-06-18 LSFA/rs.tex
esquema de RS terminado

\subsection{The Rasiowa-Sikorski lemma}
In order to formalize of the proof of the Rasiowa-Sikorski lemma in
general, a version of the Axiom of Choice ($\AC$) must be proved. This
version is known as the \emph{Principle of Dependent choices}:
\begin{quote}
  ($\DC$) Let $\rr$ be a binary relation on $A$, and $a\in A$. If
  $\forall x\in A\,  \exists y\in A\, x\rr y$, then there exists
  $f:\om\to A$ such that $f(0)=a$ and $f(n)\rr f(n+1)$ for all
  $n\in\om$.
\end{quote}

@@

>>>>

\subsection{A sequence version of $\DC$}
In order to prove the Rasiowa-Sikorski lemma in
general, a version of the Axiom of Choice ($\AC$) must be proved. This
version is known as the \emph{Principle of Dependent choices}:
\begin{quote}
  ($\DC$) Let $\rr$ be a binary relation on $A$, and $a\in A$. If
  $\forall x\in A\,  \exists y\in A\, x\rr y$, then there exists
  $f:\om\to A$ such that $f(0)=a$ and $f(n)\rr f(n+1)$ for all
  $n\in\om$.
\end{quote}

@@

\subsection{The Rasiowa-Sikorski lemma}
In order to state this Lemma, we gather the relevant hypotheses into a locale:

\begin{isabellebody}%
\isanewline
\isacommand{locale}\isamarkupfalse%
\ countable{\isacharunderscore}generic\ {\isacharequal}\ forcing{\isacharunderscore}poset\ {\isacharplus}\isanewline
\ \ \isakeyword{fixes}\ {\isasymD}\isanewline
\ \ \isakeyword{assumes}\ countable{\isacharunderscore}subs{\isacharunderscore}of{\isacharunderscore}P{\isacharcolon}\ \ {\isachardoublequoteopen}{\isasymD}\ {\isasymin}\ nat{\isasymrightarrow}Pow{\isacharparenleft}P{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{and}\ \ \ \ \ seq{\isacharunderscore}of{\isacharunderscore}denses{\isacharcolon}\ \ \ \ \ \ \ \ {\isachardoublequoteopen}{\isasymforall}n\ {\isasymin}\ nat{\isachardot}\ dense{\isacharparenleft}{\isasymD}{\isacharbackquote}n{\isacharparenright}{\isachardoublequoteclose}\isanewline
\end{isabellebody}
%
That is, $\calD$ is a sequence of dense subsets of the poset $P$. A
filter is \emph{$\calD$-generic} if it intersects every dense set in
the sequence.

\begin{isabellebody}%
\isanewline
\isacommand{definition}\isamarkupfalse%
\ {\isacharparenleft}\isakeyword{in}\ countable{\isacharunderscore}generic{\isacharparenright}\isanewline
\ \ D{\isacharunderscore}generic\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}i{\isasymRightarrow}o{\isachardoublequoteclose}\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}D{\isacharunderscore}generic{\isacharparenleft}G{\isacharparenright}\ {\isacharequal}{\isacharequal}\ filter{\isacharparenleft}G{\isacharparenright}\ {\isasymand}\ {\isacharparenleft}{\isasymforall}n{\isasymin}nat{\isachardot}{\isacharparenleft}{\isasymD}{\isacharbackquote}n{\isacharparenright}{\isasyminter}G{\isasymnoteq}{\isadigit{0}}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\end{isabellebody}

We can now state the Rasiowa-Sikorski Lemma.
\begin{isabellebody}%
\isanewline
\isacommand{theorem}\isamarkupfalse%
\ {\isacharparenleft}\isakeyword{in}\ countable{\isacharunderscore}generic{\isacharparenright}\ rasiowa{\isacharunderscore}sikorski{\isacharcolon}\isanewline
\ \ {\isachardoublequoteopen}p{\isasymin}P\ {\isasymLongrightarrow}\ {\isasymexists}G{\isachardot}\ p{\isasymin}G\ {\isasymand}\ D{\isacharunderscore}generic{\isacharparenleft}G{\isacharparenright}{\isachardoublequoteclose}\isanewline
\end{isabellebody}

The intuitive argument for the result is simple: Once $p_0=p\in P$ is
fixed, we can recursively choose $p_{n+1}$ such that 
$p_n \geq p_{n+1}\in \calD_n$, since $\calD_n$ is dense in $P$. Then
the filter generated by $\{p_n : n\in \om\}$ intersects each set in
the sequence $\{\calD_n\}_n$. This argument appeals to the sequence
version of $\DC$; we have to prove first that the relevant relation
satisfies its hypothesis:

\begin{isabellebody}%
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isacharparenleft}\isakeyword{in}\ countable{\isacharunderscore}generic{\isacharparenright}\ RS{\isacharunderscore}relation{\isacharcolon}\isanewline
\ \ \isakeyword{assumes}\isanewline
\ \ \ \ \ \ \ \ {\isadigit{1}}{\isacharcolon}\ \ {\isachardoublequoteopen}x{\isasymin}P{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isakeyword{and}\isanewline
\ \ \ \ \ \ \ \ {\isadigit{2}}{\isacharcolon}\ \ {\isachardoublequoteopen}n{\isasymin}nat{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ {\isachardoublequoteopen}{\isasymexists}y{\isasymin}P{\isachardot}\ {\isasymlangle}x{\isacharcomma}y{\isasymrangle}\ {\isasymin}\ {\isacharparenleft}{\isasymlambda}m{\isasymin}nat{\isachardot}\ {\isacharbraceleft}{\isasymlangle}x{\isacharcomma}y{\isasymrangle}{\isasymin}P{\isacharasterisk}P{\isachardot}\ {\isasymlangle}y{\isacharcomma}x{\isasymrangle}{\isasymin}leq\ {\isasymand}\ y{\isasymin}{\isasymD}{\isacharbackquote}{\isacharparenleft}pred{\isacharparenleft}m{\isacharparenright}{\isacharparenright}{\isacharbraceright}{\isacharparenright}{\isacharbackquote}n{\isachardoublequoteclose}\isanewline
\end{isabellebody}
%
Both of the proofs written in the Isar language.
